---
title: "passenger"
author: "foster"
date: "2024-06-14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
install.packages("forecast")
install.packages("tseries")
install.packages("psycho")

```

library(tidyverse)
library(forecast)
library(tseries)
library(dplyr)
library(tibble)

### load the dataset

data("AirPassengers")

ts_data <- AirPassengers

theplot1 <- autoplot(ts_data) +
  ggtitle("Monthly Totals of International Airline Passengers") +
  xlab("Year") +
  ylab("Number of Passengers") +
  theme_classic()

```{r}

plot(theplot1)
```
ggsave("~/Desktop/ts_plot.png", theplot1)

# Decompose the time series
decomposed <- decompose(ts_data)

# Plot the decomposed components (this provides the data in different categories)
# the different decomposed data results
theplot2 <- autoplot(decomposed) +
  ggtitle("Decomposed Time Series Components") +
  theme_classic()

```{r}

plot(theplot2)

```
ggsave("~/Desktop/decomposed_plot.png", theplot2)


# Fit the ARIMA model
arima_model <- auto.arima(ts_data)

# Summary of the model
summary(arima_model)

# Forecast using the ARIMA model
arima_forecast <- forecast(arima_model, h = 24)

# Plot the forecast
theplot3 <- autoplot(arima_forecast) +
  ggtitle("ARIMA Model Forecast") +
  xlab("Year") +
  ylab("Number of Passengers") +
  theme_minimal()

```{r}

plot(theplot3)

```

ggsave("~/Desktop/arima_forecast.png", theplot3)

# Fit the ETS model
ets_model <- ets(ts_data)

# Summary of the model
summary(ets_model)


# Forecast using the ETS model
ets_forecast <- forecast(ets_model, h = 24)

# Plot the forecast
theplot4 <- autoplot(ets_forecast) +
  ggtitle("ETS Model Forecast") +
  xlab("Year") +
  ylab("Number of Passengers") +
  theme_classic()


ggsave("~/Desktop/ets_forecast.png", theplot4)


# Fit the ETS model
ets_model <- ets(ts_data)

# Summary of the model
summary(ets_model)

# Forecast using the ETS model
ets_forecast <- forecast(ets_model, h = 24)


# Calculate accuracy metrics for ARIMA model
arima_accuracy <- accuracy(arima_forecast)

# Calculate accuracy metrics for ETS model
ets_accuracy <- accuracy(ets_forecast)

# Display the accuracy metrics

```{r}
print("ARIMA Model Accuracy:")
print(arima_accuracy)

```

print("ETS Model Accuracy:")
print(ets_accuracy)


### To compare which of the models is best, we look at the summary values. The smaller one ### was indicative of the model prediction power. 



